---
title: "p8105_hw2_xy2517"
author: "Xuehan Yang"
date: "2021/10/2"
output: github_document
---


```{r, message=FALSE}
library(tidyverse)
library(readxl)
```


# Problem 1

**Read and clean the Mr.Trash Wheel sheet.**

```{r,warning=FALSE}
trashwheel = 
  read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = round(sports_balls,0)) %>%  # round sports_balls to integer
  rename(weight = weight_tons, volumn = volume_cubic_yards) %>%  # change column name
  drop_na(dumpster)

head(trashwheel,3) %>% knitr::kable()
```

**Read and clean 2018 and 2019 precipitation data.**

```{r, read data seperately}
rain_2019 = 
  read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2019 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(total,month) %>% 
  mutate(year = 2019)

rain_2018 = 
  read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(total,month) %>% 
  mutate(year = 2018)
```

**Combine the rain_2018 and rain_2019 together.**

```{r, bind rows}
precipitation = bind_rows(rain_2018,rain_2019) %>% 
  mutate(month = month.name[month]) %>%  # change numeric month name to character
  relocate(year,month)

head(precipitation,3) %>% knitr::kable()
```

**Description:**

* As for trashwheel, there are total `r nrow(trashwheel)` observations, `r ncol(trashwheel)` variables including `r names(trashwheel)`. For example, there were 4.31 tons of trash on 2014-05-16, which included 1450 plastic bottles, 1820 polystyrene and so on. The median number of sports balls in a dumpster in 2019 is `r median(pull(filter(trashwheel,year==2019),sports_balls))` units.

* As for the Precipitation, there are total `r nrow(precipitation)` observations, `r ncol(precipitation)` columns which include `r names(precipitation)`. The total precipitation in 2018 is `r sum(pull(filter(precipitation,year == 2018),total))` inches. 


# Problem 2 

**Clean the data in pols-month.**

```{r, message=FALSE}
pols = read_csv(file = "./data/pols-month.csv") %>%
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"),"-" ) %>% 
  mutate(month = month.name[as.integer(month)], president = ifelse(prez_dem == 1, "dem", "gop"), year = as.integer(year)) %>% # Replace char month with full name
  select(-prez_dem,-prez_gop,-day)

head(pols,3) %>% knitr::kable()
```

**Clean the data in snp.csv.**

Separate data into year, month and day. Then delete day and change year to a 4 digit form.

```{r, message=FALSE}
snp_raw = read_csv(file = "./data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date,into = c( "month", "day", "year"),"/") %>%
  select(-day) %>% # remove day column
  mutate(year = ifelse(as.integer(year) <= 15,as.integer(year) + 2000,as.integer(year) + 1900)) # it can be run just once.
```

Arrange year and month, then replace numeric month name with full name.

```{r}
snp = snp_raw %>% 
  arrange(as.integer(year),as.integer(month)) %>% # change character into int , then arrange
  mutate(month = month.name[as.integer(month)]) %>% 
  relocate(year,month) # let year and month be the leading variable

head(snp,3) %>% knitr::kable()
```

**Tidy the unemployment data.**

```{r, message=FALSE}
unemployment = read_csv(file = "./data/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "percentage"
  ) %>% 
  mutate(month = replace(month,month == month.abb, month.name)) %>%  #replace abbreviation to full month name using Replace()
  janitor::clean_names()

head(unemployment,3) %>% knitr::kable()
```

**Join the three datasets together.**

```{r}
merge_df = left_join(pols, snp, by = c("year", "month")) %>% 
  left_join(unemployment, by = c("year", "month"))

tail(merge_df,3) %>% knitr::kable()
```

**Description**

* pols: it contains `r nrow(pols)` observations of `r ncol(pols)` variables related to the number of national politicians who are democratic or republican ranging from Jan.`r min(pull(pols,year))` to Jun.`r max(pull(pols,year))`, totally `r max(pull(pols,year))-min(pull(pols,year))` years. Key variales include `r names(pols)`. "gop" indicates republic and "dem" indicates democratic. "gov_gop", "sen_gop" and "rep_gop" means the number of republic governors, senators and representatives. "gov_dem", "sen_dem" and "rep_dem" means the number of democratic governors, senators and representatives. 

* snp: it contains `r nrow(snp)` observations of `r ncol(snp)` variables related to Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole. It ranges from Jan.`r min(pull(snp,year))` to July.`r max(pull(snp,year))`, totally `r max(pull(snp,year))-min(pull(snp,year))` years. Key variales include `r names(snp)`, among which "close" are the closing values of the S&P stock index on the associated date.

* unemployment: it contains `r nrow(unemployment)` observations of `r ncol(unemployment)` variables related to percentages of unemployment on the associated data. It ranges from Jan.`r min(pull(unemployment,year))` to June.`r max(pull(unemployment,year))`, totally `r max(pull(unemployment,year))-min(pull(unemployment,year))` years. Key variales include `r names(unemployment)`.

* merge_df: final datasets contains `r nrow(merge_df)` observations of `r ncol(merge_df)` variables ranging from Jan.`r min(pull(merge_df,year))` to July.`r max(pull(merge_df,year))`, totally `r max(pull(merge_df,year))-min(pull(merge_df,year))` years.


# Problem 3

**Tidying the popularity of baby names data.**

```{r, message=FALSE}
baby_names = read_csv(file = "./data/Popular_Baby_Names.csv", col_types = "ccccii") %>% 
  janitor::clean_names() %>% 
  mutate(gender = str_to_lower(gender), 
         ethnicity = str_to_lower(ethnicity),
         childs_first_name = str_to_lower(childs_first_name)) %>% 
  distinct() %>%  # remove distinct rows based on all column
  mutate(
    ethnicity = recode(ethnicity, 
                       "asian and paci" = "asian and pacific islander",
                       "black non hisp" = "black non hispanic",
                       "white non hisp" = "white non hispanic"
                       ))

head(baby_names,3) %>% knitr::kable()
```

**Produce a table showing the rank in popularity of the name “Olivia” as a female baby name over time.**

```{r female olivia}
baby_names %>% 
  filter(childs_first_name == "olivia", gender == "female") %>%
  select(year_of_birth, ethnicity,rank) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank
  ) %>% knitr::kable() # produce a nicer table
```

**Produce a table showing the rank in popularity of the most popular male name over time.**

```{r find the most common name among male children}
male = baby_names %>% 
  filter(gender == "male") %>%
  arrange(desc(year_of_birth),rank)

most_pop_male = pull(male,childs_first_name)[1]
```

**Display the rank of the most popular male name `r most_pop_male` with year and ethnicity.**

```{r}
baby_names %>% 
  filter(childs_first_name == most_pop_male, gender == "male") %>%
  select(year_of_birth, ethnicity,rank) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank
  ) %>% knitr::kable() # produce a nicer table
```

**Scatterplot showing the number of children with a name against the rank in popularity of the name.**

Firstly, select the target subdatasets.

```{r select the plotting dataset}
plot_df = baby_names %>% 
  filter(gender == "male", ethnicity == "white non hispanic", year_of_birth == 2016)
```

Then,draw plots between number of children with a name and rank in popularity.

```{r scatterplot}
num_popularity_plot = 
  ggplot(plot_df, aes(x = rank, y = count)) + 
  geom_point(aes(color = "Red")) +
  xlab("Rank in popularity") +
  ylab("The number of Children with this name") +
  ggtitle("Tendency of the number of children with the popularity of of name") +
  theme(plot.title = element_text(hjust = 0.5))

num_popularity_plot
```

```{r save plot}
ggsave("./plots/num_popularity_plot.png", width = 10, height = 8)
```

